Astrophysical calibration of gravitational-wave detectors

### 2015-07-03 docs/cosmic_cal.tex
Minor paper changes

% Describe the signal model (Taylor F2 etc...)
We use the post-Newtonian Taylor F2 waveform approximation for modelling the
signal whereby the \ac{GW} phase for 2 point-particles of mass $m_{1}$ and
$m_{2}$ is given by
%
\begin{equation}
\label{eq:gwphase}
  \Psi(f) = 2\pi f t_\mathrm{c} - \phi_{\mathrm{c}} - \frac{\pi}{4} +
\frac{3}{128\eta\nu^{5}}\sum\limits_{k=0}^{7}\alpha_{k}\nu^{k}.
\end{equation}
%
Here $\eta=m_{1}m_{2}/M^{2}$ is the symmetric mass-ratio with $M=m_{1}+m_{2}$
the total mass.  The post-Newtonian expansion parameter is $\nu=(\pi M
f)^{1/3}$ and the coefficients are $\alpha$.

The frequency-domain polarisation amplitudes are then defined as
%
\begin{eqnarray}\label{eq:signal}
  \gws_{+}(f) &=& \frac{1+\cos^{2}\iota}{2D_{L}}
\frac{\mathcal{M}^{5/6}}{\pi^{2/3}}
\sqrt{\frac{5}{24}}f^{-7/6}e^{-i\Psi(f)} \nonumber \\
  \gws_{\times}(f) &=&
 i\frac{\cos^{2}\iota}{D_{L}}\frac{\mathcal{M}^{5/6}}{\pi^{2/3}}
\sqrt{\frac{5}{24}}f^{-7/6}e^{-i\Psi(f)}
\end{eqnarray}
%
where the chirp-mass $\mathcal{M}$ is defined as $\mathcal{M}=M\eta^{3/5}$,
and $\iota$ is the inclination angle.  The \ac{GW} strain measured
at the $k$'th detector is then given by
%
\begin{equation}
  \label{eq:gravsig}
   \gws^{k}(f) = \big[ F_{+}^{k}(\alpha, \delta, \psi)\gws_{+}(f) +
F_{\times}^D(\alpha, \delta, \psi)\gws_{\times}(f)\big]
\end{equation}
%
where $F_{+},F_{\times}$ are the antenna response functions which are dependent
upon the polarisation angle $\psi$ and the sky position of the source
$\alpha,\delta$.

@@

\MP{Our aim is to estimate the marginal posterior probability distribution function (pdf)
for the unknown calibration scale factor. We use Bayes' theorem for which we need to define
a likelihood function and prior pdfs on the signal parameters being estimated. In this
analysis we use the Markov chain Monte Carlo code {\tt emcee} \cite{2013PASP..125..306F} to sample 
the posterior distribution.} We use a Gaussian likelihood function given by
%
%\begin{equation}
%  p(\vec{d} | \vec{\theta}, \curlH, I) \propto \exp\Bigg[ 2\Delta
%f\sum_{k>0}\frac{|\tilde{d}(f_{k}) - \tilde{h}(f_{k})|^{2}}{S(f_{k})}\Bigg]
%\end{equation}
\begin{equation}
\fl p(\bi{d} | \btheta, \curlH, I) \propto \exp{\left[4\Delta f \sum_{k=1}^{N_{\rm det}}
  \sum_{i=i_{f_{\rm low}}}^{i_{f_{\rm high}}}\frac{\Re{\left\{ \tilde{d}^{*}_{k,i} 
\gws_{k,i}(\btheta) - \frac{1}{2}\left(\gws^*_{k,i}(\btheta)\gws_{k,i}(\btheta) + 
\tilde{d}^*_{k,i}\tilde{d}_{k,i}\right) \right\}}}{S_{k,i}}\right]}
\end{equation}
% 
where $\Delta f$ is the frequency resolution of the data, $S_k$ is the $k^{\rm th}$ detector's one 
sided noise power spectral density, and the $i$ indices increment over frequency with $i_{f_{\rm 
low}}$ and $i_{f_{\rm high}}$ corresponding to the lower and upper range in frequencies used for 
our analysis, which were 20\,Hz to 1600\,Hz respectively. The source parameters, and the 
unknown scale factor, have been subsumed into a vector $\vec{\theta}$. For a network of detectors, 
the overall likelihood function becomes a product of the likelihoods for each detectors, i.e.\
\begin{equation}
  \label{eq:mult-likeli}
  p(\bi{d}| \btheta, \curlH) = \prod \limits_D p(\bi{d}^D| \vec{\theta}, \curlH)
\end{equation}
where $D = \{H,L,V\}$.

>>>>

% Describe the signal model (Taylor F2 etc...)
We use the TaylorF2 waveform approximation (see e.g.\ \cite{2009PhRvD..80h4043B} and references 
therein) with a 3.5 post-Newtonian expansion in phase for modelling our signal (in both simulations 
and signal recovery).  For \ac{BNS} systems we use non-spinning waveforms under the assumption that 
spins will be negligible. For \ac{NS-BH} system we use a spinning waveform, but in which only the 
black hole has non-negligible spin and that its rotational angular momentum is aligned with the 
systems orbital angular momentum.

For a non-spinning system the general form of the frequency-domain polarisation amplitudes are
%
\begin{eqnarray}\label{eq:signal}
  \gws_{+}(f) &\propto \frac{1+\cos^{2}\iota}{D_{L}}
\mathcal{M}^{5/6}f^{-7/6}e^{-i\Psi(f, \nu, t_\mathrm{c}, \phi_{\mathrm{c}})} \nonumber \\
  \gws_{\times}(f) &\propto
 \frac{\cos^{2}\iota}{D_{L}}\mathcal{M}^{5/6} f^{-7/6}e^{-i\Psi(f, \nu, t_\mathrm{c}, 
\phi_{\mathrm{c}})}
\end{eqnarray}
%
where the chirp-mass $\mathcal{M}$ is defined as $\mathcal{M}=M\eta^{3/5}$, with 
$\eta=m_{1}m_{2}/M^{2}$ and $M=m_{1}+m_{2}$, $\iota$ is the inclination angle, and $\nu=(\pi 
Mf)^{1/3}$.  The \ac{GW} strain measured at the $k^{\textrm{th}}$ detector is then given by
%
\begin{equation}
  \label{eq:gravsig}
   \gws^{k}(f) = \big[ F_{+}^{k}(\alpha, \delta, \psi)\gws_{+}(f) +
F_{\times}^D(\alpha, \delta, \psi)\gws_{\times}(f)\big]
\end{equation}
%
where $F_{+},F_{\times}$ are the antenna response functions which are dependent
upon the polarisation angle $\psi$ and the sky position of the source
$\alpha,\delta$.

@@

\MP{Our aim is to estimate the marginal posterior probability distribution function (pdf)
for the unknown calibration scale factor. We use Bayes' theorem for which we need to define
a likelihood function and prior pdfs on the signal parameters being estimated. In this
analysis we use the Markov chain Monte Carlo code {\tt emcee} \cite{2013PASP..125..306F} to sample 
the posterior distribution.} We use a Gaussian likelihood function given by
%
%\begin{equation}
%  p(\vec{d} | \vec{\theta}, \curlH, I) \propto \exp\Bigg[ 2\Delta
%f\sum_{k>0}\frac{|\tilde{d}(f_{k}) - \tilde{h}(f_{k})|^{2}}{S(f_{k})}\Bigg]
%\end{equation}
\begin{equation}
\fl p(\bi{d} | \btheta, \curlH, I) \propto \exp{\left[4\Delta f \sum_{k=1}^{N_{\rm det}}
  \sum_{i=i_{f_{\rm low}}}^{i_{f_{\rm high}}}\frac{\Re{\left\{ \tilde{d}^{*}_{k,i} 
\gws_{k,i}(\btheta) - \frac{1}{2}\left(\gws^*_{k,i}(\btheta)\gws_{k,i}(\btheta) + 
\tilde{d}^*_{k,i}\tilde{d}_{k,i}\right) \right\}}}{S_{k,i}}\right]}
\end{equation}
% 
where $\Delta f$ is the frequency resolution of the data, $S_k$ is the $k^{\rm th}$ detector's one 
sided noise power spectral density, and the $i$ indices increment over frequency with $i_{f_{\rm 
low}}$ and $i_{f_{\rm high}}$ corresponding to the lower and upper range in frequencies used for 
our analysis, which were 30\,Hz to 1600\,Hz respectively. The source parameters, and the 
unknown scale factor, have been subsumed into a vector $\btheta$. For a network of detectors, 
the overall likelihood function becomes a product of the likelihoods for each detectors, i.e.\
\begin{equation}
  \label{eq:mult-likeli}
  p(\bi{d}| \btheta, \curlH) = \prod \limits_D p(\bi{d}^D| \vec{\theta}, \curlH)
\end{equation}
where $D = \{H,L,V\}$.
